Waking up a sleeping rabbit:\\ On Natural-Language Sentence Generation with FF

### 2010-02-27 trunk/Papers/crisp-ff-10/experiments.tex
bin zufrieden so

Consider first the results for $d=0$, shown in Fig.~\ref{fig:runtimes}
(a); the lines end where the planner exceeded its five-minute timeout.
``Old FF'' in the figure is original FF, ``New preprocessor'' applies
only our fixes to FF's preprocess, ``New FF'' applies all fixes.
% The other two lines represent FF with the improved preprocessor,
%using the standard enforced hillclimbing strategy and the best-first
%search with helpful actions (``New FF'') respectively.  
Just the preprocessor fixes lead to a speed-up of at least $3$ orders
of magnitude, leading to practically useful runtime -- for instance,
the 24-word sentence at $n=4$ is generated in under two seconds. Note
that, by contrast, the search configuration does not make much of a
difference here. This changes in our next experiment.

Results for $d=2$ are shown in Fig.~\ref{fig:runtimes} (b).  Here the
old FF preprocessor crashed with a segmentation fault even at
$n=1$. Best-first search with helpful actions is clearly the
best-performing search strategy here.  It takes about 14 seconds for
$n=3$, but this sentence is already quite long at 29 words, and EHC
takes 3.5 minutes on the same input.  Best-first search without
helpful actions timed out at $n=3$, illustrating the importance of
that heuristic.

>>>>

Consider first the results for $d=0$, shown in Fig.~\ref{fig:runtimes}
(a); the lines end where the planner exceeded its five-minute
timeout. Notice that the vertical axis is logarithmic. Clearly, ``Old
FF'' (the original FF) is much too slow for practical use. The same is
true -- although to a less dramatic extent -- for LAMA, which already
takes three seconds to generate the relatively short sentence at $n=3$
and times out for larger instances. ``New preprocessor'' is FF with
only our fixes to FF's preprocess, ``New FF'' applies all fixes.
% The other two lines represent FF with the improved preprocessor,
%using the standard enforced hillclimbing strategy and the best-first
%search with helpful actions (``New FF'') respectively.  
Even the preprocessor fixes alone lead to a speed-up of at least $3$
orders of magnitude, leading to practically useful runtime -- for
instance, the 24-word sentence at $n=4$ is generated in under two
seconds. Note that, by contrast, the search configuration does not
make much of a difference here. This changes in our next experiment.

Results for $d=2$ are shown in Fig.~\ref{fig:runtimes} (b).  Here the
old FF preprocessor crashed with a segmentation fault even at
$n=1$. LAMA performs better, but times out beyond $n>2$.  Best-first
search with helpful actions is clearly the best-performing search
strategy here.  It takes about 14 seconds for $n=3$, but this sentence
is already quite long at 29 words, and EHC takes 3.5 minutes on the
same input.  Best-first search without helpful actions timed out at
$n=3$, illustrating the importance of that heuristic.

### 2010-02-27 trunk/Papers/crisp-ff-10/introduction.tex
bin zufrieden so

Existing planners are relatively good at solving modest instances of
the planning problems that arise in this context. However, Koller and
Petrick \shortcite{koller10:_exper_with_plann_for_natur_languag_gener}
recently showed that as the planning problem instances from Koller and
Stone \shortcite{KolSto07} scale up, off-the-shelf planners become
slow -- indeed, {\em too slow to be practically useful in NLG
  applications.} They showed this for SGplan and several variants of
FF \cite{HoffmannNebel01}; we show it herein also for LAMA
\cite{richter:etal:aaai-08}. Koller and Petrick noted that there are
major sources of inefficiency both in the search itself and in the
commonly used preprocesses; they identified several features of the
domain (e.g.\ comparatively large predicate and operator arities) that
diverge from most IPC benchmarks.
%the sheer number of
%ground instances of facts and actions that come up in this domain.

In this paper, we revisit this situation, and analyze the sources of
inefficiency in detail for FF. Based on the analysis, we modify FF and
obtain {\em dramatically better performance that, for the first time,
  makes automated planners a realistic possibility for NLG
  applications.} To achieve this quantum leap, all that is needed are
a few minor implementation fixes and basic reconfiguration of search
options. While these fixes certainly are no contribution to planning
in their own right, we believe it a valuable -- and slightly
unsettling -- lesson learned that such banalities can make the
difference between success and failure. They can hold up progress in
planning applications for years (until a planning expert comes by and
does some code profiling). We interpret this as a strong call for
automated configuration techniques, e.g.\ along the lines of Hutter et
al.\ \shortcite{HutHooStu07}.

>>>>

Existing planners are relatively good at solving modest instances of
the planning problems that arise in this context. However, Koller and
Petrick \shortcite{koller10:_exper_with_plann_for_natur_languag_gener}
recently showed that as the planning problem instances from Koller and
Stone \shortcite{KolSto07} scale up, off-the-shelf planners become
slow -- indeed, too slow to be practically useful in NLG
applications. They showed this for SGPlan and several variants of FF
\cite{HoffmannNebel01}; we show it herein also for LAMA
\cite{richter:etal:aaai-08}. Koller and Petrick noted that there are
major sources of inefficiency both in the search itself and in the
commonly used preprocesses; they identified several features of the
domain (e.g.\ comparatively large predicate and operator arities) that
diverge from most IPC benchmarks.
%the sheer number of
%ground instances of facts and actions that come up in this domain.

In this paper, we revisit this situation, and analyze the sources of
inefficiency in detail for FF. Based on the analysis, we modify FF and
obtain dramatically better performance that, for the first time, makes
automated planners a realistic possibility for NLG applications. To
achieve this quantum leap, all that is needed are a few minor
implementation fixes and basic reconfiguration of search
options. While these fixes certainly are no contribution to planning
in their own right, we consider it a valuable -- and slightly
unsettling -- lesson learned that such banalities can make the
difference between success and failure. They can hold up progress in
planning applications for years, unless users of a planner have direct
access to its developer. We interpret this as illustrative of the
importance of implementation details such as the preprocessor for the
overall runtime, and as a strong call for automated configuration
techniques, e.g.\ along the lines of Hutter et al.\
\shortcite{HutHooStu07}.
